---
title: "Boston Simulation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Boston Simulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6,
  fig.align = "center"
)
```

```{r setup}
library(SpatialModels)
library(SpatialModels)
library(ggplot2)
library(sf)
library(ggpubr)
```

```{r}
set.seed(2)
n_fine <- 30
n_coarse <- 10
```


# Boston Simulation

In this document, we explore a simulation of pixel data gathered over the Boston area.

A map of Boston is used as a reference underlying the pixels:

```{r}
boston_map <- generate_boston_map(plot = TRUE)
```

Over this map, we can create a grid of pixels representing those taken from a satellite.
These can be taken at varying angles, so we tilt them 45 degrees.

```{r}
coarse_grid <- generate_grid(boston_map,plot = TRUE,n = c(n_coarse,n_coarse),theta = -pi/4)
```

Underneath this grid, we can create a fine grid representing the raster image of pixels
which we would like to downcalse our satellite image to:

```{r}
fine_grid <- generate_grid(boston_map,plot = TRUE,n = c(n_fine,n_fine))

ggplot()+
  geom_sf(data = boston_map)+
  geom_sf(data = fine_grid,alpha = 0)+
  geom_sf(data = coarse_grid,alpha = 0)
```

In practice, there can be many satellite images over this fixed grid taken at a variety 
of angles, but for this simulation, we use only one.

Our objective is to use only the data from the **coarse** grid to reconstruct the data
from the **fine** grid. 

We expect the ground truth to have spatial correlation. Using an exponential covariance functiion $C(d) = e^{-(d/l)^2}$, we simulate a Gaussian random field on the fine grid.

```{r}
fine_grid <- generate_gaussian_random_field(boston_map,
                                                        l = 1,
                                                        n = c(n_fine,n_fine),
                                                       plot = TRUE)

color_limits <- c(-3.5,3.5) # used for plots
```


Then, using the areas of the intersections and the values, over the fine grid, we generate a coarse grid that we might expect out of a satellite image assuming no noise.


```{r}
coarse_grid <- generate_coarse_grid_value(fine_grid,coarse_grid)
ggplot()+
  geom_sf(data = boston_map)+
  geom_sf(data = coarse_grid,aes(fill = value))+
  scale_fill_continuous(limits = color_limits)+
  ggtitle("Coarse Grid Measurements")
```

Finally, we use only the values from the coarse grid and knowledge of the pixel intersections
and areas to recover the values on the fine grid.

If our method assumes no spatial correlation, i.e, the regularization is diagonal,
then we obtain

```{r}
lambda = 0.1
diagonal_covariance <- diag(dim(fine_grid)[1])
fine_grid_diagonal <- estimate_fine_grid(fine_grid,coarse_grid,diagonal_covariance,lambda)

ggplot()+
  geom_sf(data = boston_map)+
  geom_sf(data = fine_grid_diagonal,aes(fill = estimated_value))+
  scale_fill_continuous(limits = color_limits)+
  ggtitle("Estimated Ground Truth using Diagonal Regularization")
```

However, if we assume some spatial correlation, i.e, the regularization is a scaling
of the distance function used to generate the matrix, then we obtain

```{r}
covariance_matrix <- compute_distance_covariance_matrix(fine_grid,l=1)
lambda = 0.5

fine_grid_gaussian <- estimate_fine_grid(fine_grid,coarse_grid,covariance_matrix,lambda)

ggplot()+
  geom_sf(data = boston_map)+
  geom_sf(data = fine_grid_gaussian,aes(fill = estimated_value))+
  scale_fill_continuous(limits = color_limits)+
  ggtitle("Estimated ground truth using Gaussian regularization")
```





```{r, eval = F}
ground_truth <- ggplot()+
  geom_sf(data = boston_map)+
  geom_sf(data = fine_grid,aes(fill = value))+
  scale_fill_continuous(limits = color_limits)+
  ggtitle("Ground Truth")

diagonal_reg <- ggplot()+
  geom_sf(data = boston_map)+
  geom_sf(data = fine_grid_diagonal,aes(fill = estimated_value))+
  scale_fill_continuous(limits = color_limits)+
  ggtitle("Estimated Ground Truth using Diagonal Regularization")

gaussian_reg <- ggplot()+
  geom_sf(data = boston_map)+
  geom_sf(data = fine_grid_gaussian,aes(fill = estimated_value))+
  scale_fill_continuous(limits = color_limits)+
  ggtitle("Estimated Ground Truth using Gaussian Regularization")

ggarrange(ground_truth, diagonal_reg)
ggarrange(ground_truth, gaussian_reg)
```





