---
title: "grid_exploration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{grid_exploration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(SpatialModels)
library(ggplot2)
library(sf)
library(units)
```

# Boston Map and Grid Functions

We can generate a map of boston as a base:

```{r}
boston_map <- generate_boston_map(plot = TRUE)
```
Then, we can generate grid layers:

```{r}
fine_grid <- generate_grid(boston_map,plot = TRUE,n = c(20,20))
coarse_grid <- generate_grid(boston_map,plot = TRUE,n = c(10,10),theta = -pi/4)
```

We can plot them all together:

```{r}
ggplot()+
  geom_sf(data = boston_map)+
  geom_sf(data = fine_grid,alpha = 0)+
  geom_sf(data = coarse_grid,alpha = 0)
```
# Generating random field

Suppose our underyling grid should produce values correlated based on distance between centroids. We can simulate this as follows.
```{r}

gaussian_random_field <- generate_gaussian_random_field(boston_map,
                                                        l = 1,
                                                        n = c(30,30),
                                                        plot = TRUE)


```
Suppose we want to compute the area of intersecting grid slices.

```{r}
boston_map <- generate_boston_map()
fine_grid <- generate_gaussian_random_field(boston_map,
                                                        l = 1,
                                                        n = c(40,40),
                                                        plot = TRUE)
coarse_grid <- generate_grid(boston_map,n = c(10,10),theta = -pi/4)

intersection <- st_intersection(coarse_grid,fine_grid)

intersection_centroids <- sf::st_centroid(intersection)
fine_grid_containment_matrix <- sf::st_contains(fine_grid,intersection_centroids,sparse = FALSE)
coarse_grid_containment_matrix <- sf::st_contains(coarse_grid,intersection_centroids,sparse = FALSE)

coarse_grid_areas <- as.vector(st_area(coarse_grid))

intersection_areas <- compute_intersection_areas(fine_grid,coarse_grid,plot = FALSE)
fine_area_mat <- apply(fine_grid_containment_matrix,1,function(x){intersection_areas$area*x})
coarse_area_weight_mat <- apply(coarse_grid_containment_matrix,2,function(x) 1/coarse_grid_areas*x)

A <- coarse_area_weight_mat %*% fine_area_mat

coarse_grid_value <- A %*% fine_grid$value

coarse_grid$value <- coarse_grid_value

ggplot()+geom_sf(data = boston_map)+geom_sf(data = coarse_grid,aes(fill = value))

```

Experiment in recovering x:

```{r}
centroids <- st_centroid(fine_grid)
distances <- sf::st_distance(centroids,centroids)
standardized_distances <- distances/sd(distances)

covariances <- exp(-matrix(standardized_distances,nrow = 40^2,ncol = 40^2)/1)

lambda <- sqrt(0.1)

estimated_fine_values <- solve(crossprod(A)+lambda^2*crossprod(solve(covariances)))%*%crossprod(A,coarse_grid$value)
estimated_fine_values_simple <- solve(crossprod(A)+lambda^2*crossprod(diag(1600)))%*%crossprod(A,coarse_grid$value)

fine_grid$estimated_value <- estimated_fine_values
fine_grid$estimated_value_simple <- estimated_fine_values_simple

color_limits <- c(min(fine_grid$value),max(fine_grid$value))

ggplot()+geom_sf(data = boston_map)+geom_sf(data = fine_grid,aes(fill = value))+scale_fill_continuous(limits = color_limits)
ggplot()+geom_sf(data = boston_map)+geom_sf(data = fine_grid,aes(fill = estimated_value))+scale_fill_continuous(limits = color_limits)
ggplot()+geom_sf(data = boston_map)+geom_sf(data = fine_grid,aes(fill = estimated_value_simple))+scale_fill_continuous(limits = color_limits)
```


Testing coarse grid computation:

```{r}
fine_grid <- generate_gaussian_random_field(boston_map,
                                                        l = 1,
                                                        n = c(40,40),
                                                        plot = TRUE)
coarse_grid <- generate_grid(boston_map,n = c(10,10),theta = -pi/4)

coarse_grid <- generate_coarse_grid_value(fine_grid,coarse_grid)


covariance_matrix <- compute_distance_covariance_matrix(fine_grid,l=1)
lambda = 0.5

fine_grid_gaussian <- estimate_fine_grid(fine_grid,coarse_grid,covariance_matrix,lambda)

lambda = 0.1
diagonal_covariance <- diag(dim(fine_grid)[1])
fine_grid_diagonal <- estimate_fine_grid(fine_grid,coarse_grid,diagonal_covariance,lambda)

```

```{r}

color_limits <- c(min(fine_grid$value),max(fine_grid$value))


ggplot()+
  geom_sf(data = boston_map)

ggplot()+
  geom_sf(data = boston_map)+
  geom_sf(data = coarse_grid,alpha = 0)

ggplot()+
  geom_sf(data = boston_map)+
  geom_sf(data = fine_grid,alpha = 0)

ggplot()+
  geom_sf(data = boston_map)+
  geom_sf(data = fine_grid,alpha = 0)+
  geom_sf(data = coarse_grid,alpha = 0)


ggplot()+
  geom_sf(data = boston_map)+
  geom_sf(data = fine_grid,aes(fill = value))+
  scale_fill_continuous(limits = color_limits)

ggplot()+
  geom_sf(data = boston_map)+
  geom_sf(data = coarse_grid,aes(fill = value))+
  scale_fill_continuous(limits = color_limits)

ggplot()+
  geom_sf(data = boston_map)+
  geom_sf(data = fine_grid_diagonal,aes(fill = estimated_value))+
  scale_fill_continuous(limits = color_limits)

ggplot()+
  geom_sf(data = boston_map)+
  geom_sf(data = fine_grid_gaussian,aes(fill = estimated_value))+
  scale_fill_continuous(limits = color_limits)

ggplot()+
  geom_sf(data = boston_map)+
  geom_sf(data = fine_grid,aes(fill = value))+
  scale_fill_continuous(limits = color_limits)
```

